---
title: "homework4"
author: "Huang Xubin 3180102999"
date: "2021/7/15"
output: pdf_document
---

# 1

```{r echo = FALSE, message = FALSE}
library(ggplot2)
library(tidyverse)
```

```{r message = FALSE}
# data about the month:
ckm_nodes <- read_csv("data\\ckm_nodes.csv")

# data about the contacts:
ckm_network <- read.table("data\\ckm_network.dat")

# a Boolean vector that shows whether the doctor got the adopted month 
# recorded:
dct.bool.got_data <- !is.na(ckm_nodes$adoption_date)

# the vector of indexes of those doctors:
dct.idx.got_data <- which(dct.bool.got_data)

# cut off all useless rows in month data and rows and cols of those 
# in contacts data:
ckm_nodes <- ckm_nodes[dct.idx.got_data,]
ckm_network <- ckm_network[dct.idx.got_data, dct.idx.got_data]
dim(ckm_nodes)
dim(ckm_network)
```

# 2

```{r}
# number of rows after cleaning the data:
dct.n <- nrow(ckm_nodes)

# make a combination of doctor number and month:
dct.tbl.a <- ckm_nodes %>% mutate(begin = 1, doctor = 1 : nrow(ckm_nodes)) %>% 
  dplyr::select(doctor, month = adoption_date, begin)

# develop a table of indexes of each obs:
dct.tbl <- data.frame(doctor = rep(1 : dct.n, each = 17), 
                      month = rep(1 : 17, times = 125))

# do one left join operation to form a one-hot vector by hand:
dct.tbl <- dplyr::left_join(dct.tbl, dct.tbl.a, by = c("doctor", "month"))
dct.tbl$begin[is.na(dct.tbl$begin)] <- 0

# whether that doctor had begun before that month:
dct.tbl <- dct.tbl %>% group_by(doctor) %>% 
  mutate(begin_before = (cumsum(begin) - begin)) %>% 
  ungroup()

# table of the number of contacts of each doctor in each month that begins adopting:
invisible(dct.tbl.contacts_each_month <- data.frame(
  doctor = rep(1 : dct.n, times = rowSums(ckm_network)), 
  month = ckm_nodes$adoption_date[
    unlist(apply(as.matrix(ckm_network), 1,
                 function(e){return(which(as.logical(e)))}))]) %>% 
  group_by(doctor, month) %>% summarise(contacts.begin = n()))

# left join into the ultra table:
invisible(dct.tbl <- dplyr::left_join(dct.tbl, dct.tbl.contacts_each_month))
dct.tbl$contacts.begin[is.na(dct.tbl$contacts.begin)] <- 0

# adding the two last columns into the ultra table:
dct.tbl <- dct.tbl %>% group_by(doctor) %>% 
  mutate(contacts.begin_before = cumsum(contacts.begin) - contacts.begin, 
         contacts.begin_in_or_before = cumsum(contacts.begin)) %>% 
  ungroup() %>% dplyr::select(-contacts.begin)
head(dct.tbl)
tail(dct.tbl)
```

where 6 columns of which 2 columns are for **identifying each observation** and 4 columns left are just for the data required, and 2125 rows because **125** (number of doctor) **plus 17** (max adopted time) **equals 2125**.

# 3

## a

The reason why there should be no more than 21 entries of k is just there is no one who got more than 20 contacts, and by counting up all cases plus the 0 case gives the number of 21.

```{r}
# there is no one who got more than 20 contacts:
max(rowSums(ckm_network))
```

## b

```{r fig.width = 7, fig.height = 2.0}
# generating p_k:
dct.p.dmnt <- dct.tbl %>% filter(begin_before == 0) %>% 
  group_by(contacts.begin_before) %>% summarise(dominator = n())
dct.p.nmrt <- dct.tbl %>% filter(begin == 1) %>% 
  group_by(contacts.begin_before) %>% summarise(numerator = n())
invisible(dct.p <- dplyr::full_join(dct.p.dmnt, dct.p.nmrt) %>% 
  mutate(pr = numerator / dominator))
dct.p[is.na(dct.p)] <- 0
dct.p

# plot:
dct.p %>% ggplot() + geom_line(aes(x = contacts.begin_before, y = pr))
```

## c

```{r fig.width = 7, fig.height = 3}
# generating q_k:
dct.q.dmnt <- dct.tbl %>% filter(begin_before == 0) %>% 
  group_by(contacts.begin_in_or_before) %>% summarise(dominator = n())
dct.q.nmrt <- dct.tbl %>% filter(begin == 1) %>% 
  group_by(contacts.begin_in_or_before) %>% summarise(numerator = n())
invisible(dct.q <- dplyr::full_join(dct.q.dmnt, dct.q.nmrt) %>% 
  mutate(pr = numerator / dominator))
dct.q[is.na(dct.q)] <- 0
dct.q

# plot:
dct.q %>% ggplot() + geom_line(aes(x = contacts.begin_in_or_before, y = pr))
```
\newpage

# 4

## a

```{r}
# estimation via least squares:
f1 <- function(a, b, X = dct.p$contacts.begin_before){
  return(a + b * X)
}
f2 <- function(a, b, X = dct.p$contacts.begin_before){
  return(exp(a + b * X) / (1 + exp(a + b * X)))
}
ls <- function(params, f, Y = dct.p$pr){
  return(mean((Y - f(params[1], params[2]))^2))
}
p1 <- c(0.03284105 , 0.03459169)
result_1 <- nlm(ls, p1, f1)$estimate
result_1
```
where the parameters estimated are $(a,\ b) = (0.03284105,\ 0.03459169)$.

## b

Suppose $b > 0$, then according to the derivative of the model w.r.t. k, the image of this model is at first a steep curve right up to the top (probability of 1) and soon becomes a straight (horizontal) line heading to the right end, which indicates the impact of adding 1 contacts is at first a factor to consider and soon being nothing to consider about.

```{r}
p2 <- c(-3.6872181, 0.3834268)
result_2 <- nlm(ls, p2, f2)$estimate
result_2
```
where the parameters estimated are $(a,\ b) = (-3.6872181,\ 0.3834268)$.


## c

```{r fig.width = 7, fig.height = 4}
interval.left <- min(dct.p$contacts.begin_before)
interval.right <- max(dct.p$contacts.begin_before)
X <- seq(interval.left, interval.right, (interval.right - interval.left) / 10000)

# plot all the 3 plots:
ggplot() + geom_line(aes(x = dct.p$contacts.begin_before, y = dct.p$pr)) + 
  geom_line(aes(x = X, y = f1(result_1[1], result_1[2], X)), colour = 'red') + 
  geom_line(aes(x = X, y = f2(result_2[1], result_2[2], X)), colour = 'blue') + 
  labs(x = "k", y = "probability", title = "all 3 plots w.r.t. pk") + 
  geom_text(data = data.frame(x = 5, y = 0.21), 
            aes(x, y, label = "a + bk"), hjust = 0, vjust = 0) + 
  geom_text(data = data.frame(x = 2.5, y = 0.07), 
            aes(x, y, label = "exp(a + bk)/(1 + exp(a + bk))"), hjust = 0, vjust = 0)
```

Clearly the exponential model fits more of the origin data according to the image above.










